```{r}

```

---
title: "Final Project Data"
author: "Tan Vo"
format: html
editor: visual
embed-resources: true
---

```{r}
#| include: false
library(tidyverse)
library(skimr)
library(dplyr)
```

```{r}
#| message: false
life_expectancy <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-12-05/life_expectancy.csv')
life_expectancy_different_ages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-12-05/life_expectancy_different_ages.csv')
life_expectancy_female_male <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-12-05/life_expectancy_female_male.csv')
skim(life_expectancy)
skim(life_expectancy_different_ages)
skim(life_expectancy_female_male)
```

-   [**DATA EVALUATION**]{.underline}

Is the data in the right format for those mappings? - Yes.

Determine if the variables you have are the appropriate variable type. - Yes, it is appropriate.

Does something you want to use as a numeric variable look like a character string? - No.

See if there are any missing values, and if there are, if they are coded correctly. - No missing value.

Are the variable names usable? - Yes.

Do they have spaces in them that will make it hard to refer to them in R? - No.

For categorical variables, consider the categories. Are they readable? - Yes.

Are they understandable phrases or cryptic codes? - They are understandable phrases.

Are there the right number of categories, or will you want to reduce them? - They are correct but might consider to group them into different groups.

-   [**DATA WRANGLING**]{.underline}

1.  **Average Life Expectancy from 1950 to 2021**

```{r}
avg_life_expectancy <- life_expectancy |>
  filter(Year >= 1950 & Year <= 2021, !is.na(Code), Code != "OWID_WRL") |>
  group_by(Entity, Code) |>
  summarize(AvgLifeExpectancy = mean(LifeExpectancy, na.rm = TRUE)) |>
  ungroup()
```

Export the filtered data as a CSV file

```{r}
write_csv(avg_life_expectancy, "avg_life_expectancy.csv")
```

2.  **Average Life Expectancy Difference between Female & Male from 1950 to 2021**

```{r}
avg_life_expectancy_diff_female_male <- life_expectancy_female_male |>
  filter(Year >= 1950 & Year <= 2021, !is.na(Code), Code != "OWID_WRL") |>
  group_by(Entity, Code) |>
  summarize(AvgLifeExpectancyDiffFM = mean(LifeExpectancyDiffFM, na.rm = TRUE)) |>
  ungroup()
```

Export the filtered data as a CSV file

```{r}
write_csv(avg_life_expectancy_diff_female_male, "avg_life_expectancy_diff_female_male.csv")
```
